A Bioconductor workflow for processing and analysing spatial proteomics data

### 2018-04-16 replies-to-reviewers.tex

To address the above comment on suitability we have added a few additional points on the challenges of using clustering for this type of data. \\

We generally find supervised learning more suited to the task of protein localisation prediction in which we use high-quality curated marker proteins to build a classifier, instead of using an entirely unsupervised approach to look for clusters and then look for enrichment of organelles and complexes. In the latter we do not make good use of valuable prior knowledge, and in our experience unsupervised clustering can be extremely difficult due to (i) the loose definition of what constitutes a cluster (for example whether it is defined by the quantitative data or the localisation information), (ii) the influence of the algorithm assumption on the cluster identification (for example parametric or non-parametric) and (iii) poor estimates of the number of clusters that may appear in the data. \newline \newline

\textbf{Minor comments:} \\
\textcolor{gray} {I was not able to naively reproduce the workflow from the R commands in the article due to an error installing pRolocdata on a Windows machine. On OS X it was smooth.}

We didn't experience any problems with installing \texttt{pRolocdata} on Windows. If you re-try the installation and please let us know if you still have any issues by opening an issue (https://github.com/lgatto/pRolocdata/issues) or by posting this issue on the Bioconductor Support site (https://support.bioconductor.org). \\

@@

\small{
\begin{lstlisting}[language=R] 
> library(pRoloc)
> library(pRolocdata)
> data(itzhak2016stcSILAC)
> ## 6 combined replicates of 5 fractions each
> dim(itzhak2016stcSILAC)
[1] 5265   30
> plot2D(itzhak2016stcSILAC)
\end{lstlisting}}

If there were only very few fractions (say cytosol, nucleus and plasma membrane), the infrastructure could still be used, although for less benefits. \\

\textcolor{gray} {How do the authors recommend collapsing replicates? This could be covered in the section dedicated to the Compare function. Two replicates will (almost) never agree 100\% so how are discrepancies handled?}

Currently, we recommend to visualise different replicates on their own, to confirm that they are of sufficient quality, and then combine them, retaining the proteins that have been quantified over all replicatedd experiments. This allows to obtain localisation information over all replicated data. We however do not explicitly assess the variability using this approach. This could be done by analysing replicated independently and then compare the coherence of the classification results. Proteins that are only observed in some replicates could be rescued by repeating the analysis using only the relevant (possibly unique) replicate(s). \\ \newline
